Big Data with High Performance Computing (HPC)

### 2016-04-10 Tex_Files/paper.tex
Final Draft - Parallel database and IO

@@

One of such specialized Deep Learning accelerators is the one introduced by Chen \textit{et al.}, namely DianNao \cite{chen2014diannao}. DianNao comprises two major group of components. First, buffers for retrieving and holding inputs and outputs of neurons and synapses. Second, a Neural Functional Unit (NFU), which performs computations to calculate outputs of neurons. These calculations are mainly matrix multiplication of weights and input in the first stage, summation of the results of first stage in the second stage, and applying an activation function in the third stage. According to the architecture of network and the characteristics of layers some other types of operation is also embedded such as convolution, averaging, etc. Figure \ref{fig:diannao_arch} depicts a simplified block diagram of the DianNao accelerator.
\begin{figure}[h]
	\includegraphics[scale=0.23]{./images/diannao.png}
	\centering
	\caption{Block diagram of the DianNao accelerator\cite{chen2014dadiannao}.}
	\label{fig:diannao_arch}
\end{figure}
To evaluate the performance of DianNao, Chen et al. structures two neural networks with similar architecture, but one with GPU and the other one with DianNao acceleration. They also implemented the same network using CPU only without any acceleration. Figure \ref{fig:diannao_speedup} compares the speedup of GPU over both DianNao and CPU.
\begin{figure}[h]
	\includegraphics[width=\textwidth]{./images/diannao_results.png}
	\centering
	\caption{Speedup of CPU/GPU and CPU/DianNao for various layer types of network\cite{chen2014dadiannao}.}
	\label{fig:diannao_speedup}
\end{figure}

We can see from figure \ref{fig:diannao_speedup} that DianNao gives speedups up to 100 for some of the layers. However, in some other layers, mostly convolutional layers with private kernels and classification layers, which are common in both Deep Neural Networks (DNN) and Convolutional Neural Networks (CNN), GPUs may be preferred over DianNao. Authors believe that memory bandwidth is the major reason that DianNao performs worse than GPU in such layers. It should be noted that both convolutional and classification layers share the property that they have high number of synaptic connections in the range of millions. As DianNao takes up only about 53\% space of the a equivalent GPU (K20M), which is used for the sake of this comparison, it has shown the potential to compete with GPUs.
\subsection*{Dadiannao, the machine learning supercomputer}

While the authors name Dadiannao the machine learning supercomputer, it seems that design of Dadiannao was more focused on DNNs and CNNs. With that in mind, Dadiannao is capable of holding large models of even up to tens of GBs in size in a multi node structure.\\

Earlier we sat that, memory is thought to be the bottleneck of DianNao in convolutional layers with private kernel and classifier layers. The design of Dadiannao as a supercomputer aims to address these problems. First, in order to minimize data movement cost, synapses are placed near to the neurons consuming their data in a fully distributed architecture with no main memory unit. Secondly, nodes are biased towards storing data rather than computation. Thirdly, since the number of synaptic connections is orders of magnitude bigger than neurons,  instead of sending synaptic weights to neurons, neuron values is sent to synapses. This saves a lot in inter node communication bandwidth. Fourthly, intra-node, the storage units are broken into many storage units to increase the bandwidth.\\

@@

\begin{figure}[h]
	\includegraphics[scale=0.42]{./images/NFU_diagram.png}
	\centering
	\caption{Block diagram of a node within Dadiannao architecture \cite{chen2014dadiannao}.}
	\label{fig:NFU_Diagram}
\end{figure}
It should be noted that although SRAMs are generally preferred over other types of memory such as DRAM and eDRAM due to the higher speed, their size is limiting factor for them. A typical eDRAM can accommodate more than $2.5\times$ data compared to a SRAM. It should be noted that more storage density is required if parameters of big models are to be stored within chip to avoid external memory access. As a result, eDRAM is selected over SRAM in Dadiannao's proposed design. Refuting the memory bandwidth limitations for for computational units allows for scaling since more inputs and outputs to each neuron can be calculated simultaneously. However, eDRAMs suffer from some drawbacks compared to SRAMs namely, slower response, periodical refreshing requirement, and destructive read that may result in inaccurate values.\\

In order to exploit high bandwidth within nodes, a tile fashion layout is implemented as shown in figure \ref{fig:tiles}. In this design output neurons are assigned to various tile units. This enables NFUs to process data from multiple neurons in parallel.
\begin{figure}[h]
	\includegraphics[scale=0.42]{./images/tiles.png}
	\centering
	\caption{Left: Tile implementation of a node. Right: Layout of unit inside a tile. \cite{chen2014dadiannao}.}
	\label{fig:tiles}
\end{figure}

Tiles in figure \ref{fig:tiles} are connected to two eDRAM reservoirs in the middle, one for input neurons from tiles and one for output neurons. It should be noted that even though there are many tiles and NFUs in a node, the number of neurons in a typical deep learning architecture is much more to be a one to one relationship between NFUs and neurons. As a result, each NFU has the job of calculating many neuron outputs.

\subsection*{Experimental Results}
In this part we present the performance and energy consumption Dadiannao. In figure \ref{fig:diannao_speedup}, performance of Dadiannao in the training mode is compared against its NVIDIA K20M GPU implementation counterpart.
\begin{figure}[h]
	\includegraphics[scale=0.45]{./images/dadiannao_speedup.png}
	\centering
	\caption{Speedup of Dadiannao/GPU for various types of layers. CONV1 and fullNN layers require at least 4 nodes. CONV3* and CONV4* need at least 36 nodes to operate\cite{chen2014dadiannao}.}
	\label{fig:dadiannao_speedup}
\end{figure}

Figure \ref{fig:dadiannao_speedup} shows that Dadiannao could reach a speedup of more than 100 over the GPU system in almost all types of layers. The speedup is even higher up to more than 2000 for the CONV1 and CONV2 layers. It should be noted that the size of CONV3* and CONV4* layers (which have private kernels) is relatively large in the range of GB. As a result, the performance of these layers can only be measured of on a more than 36-node arrangement. The fullNN layer, which is a fully connected layer also accounts for more than 50M synaptic connections and hence cannot be implemented on a system with only 1 node. Generally speaking, average speedup of $21.38\times,$ $79.81\times,$ $216.72\times,$and $450.65\times$ is reported for Dadiannao over GPU for 1-node, 4-node, 16-node, and 64-node configurations, respectively. However, the speedup and scaling of various types of layers are wildly different due to the inherent characteristics of each type of layer.\\

Similarly, the performance of Dadiannao against GPU baseline in evaluated from energy point of view. Figures \ref{fig:dadiannao_training} and \ref{fig:dadiannao_inference} depict these performance analysis for both training and inference modes. In the training mode the energy reduction rates against the GPU baseline are $172.39\times,$ $180.42\times,$ $142.59\times,$ and $66.94\times$ for the 1-node, 4-node, 16-node, and 64-node respectively. The energy reduction number reported in the inference mode is even higher. In the inference mode, the energy reduction for the 1-node, 4-node, 16-node, and 64-node configuration is $330.56\times,$ $323.74\times,$ $270.04\times$, and $150.31\times,$ respectively.

\begin{figure}[!htb]
	\includegraphics[scale=0.45]{./images/dadiannao_training.png}
	\centering
	\caption{Comparison of energy reduction of Dadiannao against GPU in the training mode\cite{chen2014dadiannao}.}
	\label{fig:dadiannao_training}
\end{figure}
\begin{figure}[!htb]
	\includegraphics[scale=0.45]{./images/dadiannao_inference.png}
	\centering
	\caption{Comparison of energy reduction of Dadiannao against GPU in the inference mode\cite{chen2014dadiannao}.}
	\label{fig:dadiannao_inference}
\end{figure}

\section*{Conclusion and Outlook}
Machine learning and deep learning  have already become a part of our lives that we even do not notice them anymore. From search engines, voice recognition systems in our phones, and self-driving cars, all the way to commercial and financial analytics, they all benefit from machine learning and deep learning. With the help of fast CPUs and GPUs, training of models within machine learning and deep learning frame encountered a huge boost in the recent past. However, it seems that the future attempts to guarantee the the scalability of such systems should be focused on designing specialized software and hardware frameworks tailored to these applications such as project Adam and Dadiannao supercomputer.\\
\subsection{High Performance Graph Processing}
Graphs are abstract mathematical concepts that have numerous applications in science and engineering. Depending on the application and context they are being adopted in, they may have different types such a directed or undirected, cyclic or acyclic, finite or infinite, etc.\\
\begin{figure}[!ht]
	\includegraphics[scale=0.38]{./images/BSP_model.png}
	\centering
	\caption{Execution mechanism of BSP programming model \cite{sakr2013processing}.}
	\label{fig:BSP}
\end{figure}
With the explosion of data over the past decade, graphs have dictated themselves as tools with many interesting properties. Nowadays, graphs are an indispensable component of large data modeling systems. Graph databases are being more and more favored over relational databases with focus on relations as connections indicate foreign-keys in the traditional relational databases. Among many interesting inherent properties of graphs, their scalability is probably the most important distinguishing factor.\\

@@

Each vertex has the option of becoming active or inactive during each iteration. If a vertex does not receive a message it can vote to halt and become inactive. Similarly, an inactive can become active upon arrival of a message. The overall execution continues until no vertex has a message to send. Figure \ref{fig:max_super} illustrates an example of finding the maximum value in a graph.
\begin{figure}[!ht]
	\includegraphics[scale=0.38]{./images/max_superstep.png}
	\centering
	\caption{Example of computing maximum vertex value. In the first superstep values for each vertex are sent to the neighbor. If the received value is higher than the current value, current value is updated. Otherwise, current value is hold.  \cite{sakr2013processing}.}
	\label{fig:max_super}
\end{figure}

%%%%%%% KRUNAL ADDED
\subsection{Parallel Databases and I/O}
Although the idea of parallel database system was refuted by a paper prophesying demise of database machine in 1983 but that did not stop the adoption of parallel database. Companies like Teradata, Tandem and many other startup companies pioneered a path for successful parallel database systems. The success of parallel database is accorded to wide adoption of relational model which favors the idea of parallel dataflow graph leading to pipelined parallelism and partitioned parallelism. Ten years later a paper from Dewitt explains how parallel database system would be the future of high performance database systems. With the advances in high speed interconnect network, the client-server operating system gave much higher performance over price than its mainframe equivalent.\\

@@

Although, there is a heated argument about the performance benefits of using a Map Reduce system against the parallel DBMS system, former has been widely adopted because it offers better performance with complex queries on unstructured data. High performance computing systems have high speed network, parallel file systems and diskless nodes while Big data analytics research has evolved from an increasing demand for extracting knowledge from large data set on a big and small cluster with low speed network with each node having its own local file system and a disk. For example, Lustre file system is widely used in HPC environment whereas Hadoop's HDFS is popular for big data analytic applications. The main focus of HPC machine is optimize and exploit the resources, with a parallel file system. During the past decade, several attempts are made to bridge the gap between HPC and Big data analytics giving rise to new terms called High Performance Data Analysis. In the next section, we discuss about some of these research that has been propelled towards having an amalgamation of these two giants.\\

\subsubsection{Project SCOPE: Parallel databases meet Map/Reduce}
Structure Computation Optimization for Parallel Execution (Scope) is a distributed computing system, that incorporates best characteristics of traditional parallel database system and map reduce execution engine like scalability and robust implementation. It is being used in many of the online service from Microsoft like Bing. A physical execution plan containing a directed acyclic graph are crafted from a script by a Cascade framework based optimizer.  Like MapReduce, a Job Manager provides fault tolerance and recovery and at the same time a declarative scripting language similar to SQL that provides abstraction and highly extensibility through user defined functions and relational operations.\\

\subsubsection{Platform Architecture:}Structure Computation Optimization for Parallel Execution (Scope) is a distributed computing system, that incorporates best characteristics of traditional parallel database system and map reduce execution engine like scalability and robust implementation. It is being used in many of the online service from Microsoft like Bing. A physical execution plan containing a directed acyclic graph are crafted from a script by a Cascade framework based optimizer.  Like MapReduce, a Job Manager provides fault tolerance and recovery and at the same time a declarative scripting language similar to SQL that provides abstraction and highly extensibility through user defined functions and relational operations.

\begin{figure}[!htb]
	\includegraphics[scale=1]{./images/krunal1}
	\centering
	\caption{Architecture of cosmos platform }
	\label{fig:architecture_of_cosmos}
\end{figure}
Writes are always append-only and serialization is used for concurrent writes. The storage system consist of a physical partitions called extends, which are the unit of space allocation and replication, typically of few 100 MB size. Application extends, often compressed and decompressed at client side, contains application-defined records and define a block boundary. Local as well as remote read and write are enabled through high-bandwidth network. Directed acyclic graph (DAG) are used for modelling a query plan. Job submission and management, job queue monitoring, interoperability, job status tracking, error reporting and transferring data in and out of COSMOS are provided through an interface from frontend services.\\

\subsection*{Data Representations-Streams}
It supports both structure and unstructured stream. Unstructured stream are defined as a logical sequence of bytes that requires an extractor for users to interpret it. An extractor defines the schema and iterator interface for transforming the data. It also provides an outputters that writes output in an unstructured streams. Structured stream are handled by build-in function that allow an unvarying access time by storing data with into different schemas which are self-contained, that is, with rich metadata information such as schemas, structural properties and access method.\\

Usually these data are horizontally portioned with different schemes like hash and range partitioning. Each structural stream can consist of several physical extents which allows effective replication and fault tolerance. Data is stored based on store affinity that assigs affinity id, with each affinity group having same id. Storage priority is given first at same machine level, if the storage is full then at same rank and then at next close rank.  Another optimization to improve performance is stream references, in this technique the output stream is referenced with another corresponding partition based on affinity id based on the storage priority shown above. Some other features are Indexing for random access, column group for vertical partitioning.  \\

\subsubsection{Query Language for Scope}
SCOPE query language resembles SQL and has many SQL-like extensions that enable users to separate application logic from underline implementation on a distributed system. For unstructured stream, it provides personalized commands for EXTRACT and OUTPUT, for reading from source and writing data to a sink. Structure stream removes the constraints of specifying columns because it is already contained in the stream metadata during compilation. Reference clause can be used to specify affinity of output to another stream. REDUCE command can be replaced by adding an optional PRESORT and PRODUCE clause. Figure \ref{fig:compilation_of_a_scope} shows the all steps of transformation for a SCOPE script before executing in the cluster.

\begin{figure}[!htb]
	\includegraphics[scale=1]{./images/krunal2}
	\centering
	\caption{Compilation of a SCOPE script.}
	\label{fig:compilation_of_a_scope}
\end{figure}

\subsubsection{Code Generation and Runtime Engine}
SCOPE runtime engine supports all the relational operators along with build in and optimized UDOs like DefaultTextExtractor and DefaultTextOutputter. Simple functions like open, next and close are performed by operator which inherently supports pipelining of these operators.  The runtime engine, after optimization, generates a physical execution tree by doing these tasks:
\begin{enumerate}
	\item The whole plan is analyzed to produce corresponding codes for operator based on the internal operator template.
	\item  Output tree is partitioned at exchanges and spool operators to generate a super vertex that contains a series of physical exchanges.
	\item Break large super vertex into smaller ones to avoid skew delay due to longer execution time.
	\item An assembly is made by compiling all the generated code into it and the entry functions for each super vertex is written into super vertex definition.
	\item Each super-vertex is schedule to execute on a single machine.
	\item Like traditional db system, a queue of additional super vertices are maintained through admission control technique when all resources are allocated.
\end{enumerate}

@@

\subsubsection{Execusion Model}
Similar to MapReduce, SCOPE implements a pull execution model. Super vertex, reads data either locally or remotely, are executed in a pipelined fashion with an end result written to a local disk and waiting for the next vertex to pull data. It contrast with the execution model of parallel database which transfer intermediate data directly without any disk access. However, pull execution model has an advantage that both parties involved in data exchange need not run simultaneously and the data can be recovered from disk on failure. Since shuffling of data is the most expensive part considering the network overhead, SCOPE runtime provided many partitioning and merge operators.  It supports hash, range and index-based partitioning, with number of partitions predetermined or determined based on intermediate data size during runtime.

\subsubsection{Job Scheduling}
SCOPE's job manager is an adaption from Dryad, which takes care of job graph and resource scheduling on clusters. DAG of super vertices is scheduled to execute separately on different machines. Vertices are classified into different stages by Job manager and all the vertices in a single stage perform same task.  Data storage and computation can be performed by all the machine.\\
\begin{figure}[!htb]
	\includegraphics[scale=1]{./images/krunal3}
	\centering
	\caption{SCOPE job scheduling }
	\label{fig:scope_job_sched}
\end{figure}
Local system resources are maintained by processing node services and it schedules execution of local vertices along with monitoring state of a job. Based on available resources and priorities it generates process on behalf of Job managers and maintains a local queues of vertex execution request from different Job manager. It stores the assembly of vertex in a cache and uses it if the vertex is executed again. An initial DAF is formed, when the job begins, to oversee relationship among different vertex. According to execution results and variation in system environment, the job graph is dynamically adapts itself. Resource availability and vertex priority defines the actual vertex scheduling order and it's the task of Job Manager to notify any vertex execution failure.\\

\subsubsection{Adaptive Hybrid OLAP Architecture}
Since last decade, co-processor technology has evolved rapid due to huge popularity of high resolution and graphic intensive games. These co-processors are very good at processing computationally intensive tasks and since data analytics usually requires a single operation on a multiple datasets (SIMD), it's turns out to be perfect candidate. In the most of the research, the CPU ( device) acts as a supervisor and GPU (host) carries out all the task which leads to inefficient use of device resources. \\

An adaptive hybrid OLAP (On-line Analytical Processing) architecture is proposed in XXXXXXPAPER that uses prediction model that can efficiently distribute and process data on CPU and GPU based on their memory subsystem and memory access pattern. OLAP has a fundamental structure of a data cube which consist of multi-dimensional data that enables user to explore relation amongst different aspects of data. Data is partitioned into two types of OLAP cubes for processing on CPU and GPU:\\
\begin{enumerate}
	\item Multidimensional OLAP (MOLAP): These cubes contains preprocessed data that is sparse and huge in size. It is coherent to CPU's memory system.
	\item Relational OLAP (ROLAP): It contains data that is manipulated to appear like traditional OLAP with slicing and dicing functionality, which is suitable to be processed on GPU architecture.
\end{enumerate}

>>>>

\usepackage{amsmath}
\usepackage{algpseudocode}
\usepackage{algorithm}
%\usepackage[showframe]{geometry}%<--- only to show the page structure

@@

One of such specialized Deep Learning accelerators is the one introduced by Chen \textit{et al.}, namely DianNao \cite{chen2014diannao}. DianNao comprises two major group of components. First, buffers for retrieving and holding inputs and outputs of neurons and synapses. Second, a Neural Functional Unit (NFU), which performs computations to calculate outputs of neurons. These calculations are mainly matrix multiplication of weights and input in the first stage, summation of the results of first stage in the second stage, and applying an activation function in the third stage. According to the architecture of network and the characteristics of layers some other types of operation is also embedded such as convolution, averaging, etc. Figure \ref{fig:diannao_arch} depicts a simplified block diagram of the DianNao accelerator.
\begin{figure}[h]
	\includegraphics[scale=0.23]{./images/diannao.png}
	\centering
	\caption{Block diagram of the DianNao accelerator\cite{chen2014dadiannao}.}
	\label{fig:diannao_arch}
\end{figure}
To evaluate the performance of DianNao, Chen et al. structures two neural networks with similar architecture, but one with GPU (NVIDIA K20M) and the other one with DianNao acceleration. They also implemented the same network using CPU only without any acceleration. Figure \ref{fig:diannao_speedup} compares the speedup of GPU over both DianNao and CPU. It should be noted that the code for the GPU is developed using CUDA language and the CPU code is in C++ for different types of layers in the network.
\begin{figure}[h]
	\includegraphics[width=\textwidth]{./images/diannao_results.png}
	\centering
	\caption{Speedup of CPU/GPU and CPU/DianNao for various layer types of network\cite{chen2014dadiannao}.}
	\label{fig:diannao_speedup}
\end{figure}

Figure \ref{fig:diannao_speedup} portrays that GPU has a speedup of around 50x to 100x for most layers. DianNao gives speedups up to 100 for some of the layers. However, in some other layers, mostly convolutional layers with private kernels and classification layers, which are common in both Deep Neural Networks (DNN) and Convolutional Neural Networks (CNN), GPUs may be preferred over DianNao. Authors believe that memory bandwidth is the major reason that DianNao performs worse than GPU in such layers. It should be noted that both convolutional and classification layers share the property that they have high number of synaptic connections in the range of millions. As DianNao takes up only about 53\% space of the a equivalent GPU, which is used for the sake of this comparison, it has shown the potential to compete with GPUs.
\subsection*{DaDianNao, the Machine Learning supercomputer}

DaDianNao is designed to serve as a Machine Learning supercomputer tailored for DNNs and CNNs. 
Dadiannao has a performance better or on a par with GPU options while taking up less area at lower costs. Besides, it can hold large models of even up to tens of GBs in size in a multi node structure.\\

As explained earlier, memory is thought to be the bottleneck of DianNao in convolutional layers with private kernel and classifier layers. The design of Dadiannao as a supercomputer aims to address these problems. First, in order to minimize data movement cost, synapses are placed near to the neurons consuming their data in a fully distributed architecture with no main memory unit. Secondly, nodes are designed so they tend to store instead of computation. Thirdly, since the number of synaptic connections is way more than the number of neurons,  instead of sending synaptic weights to neurons, neuron values are sent to synapses. This saves a lot in external communication bandwidth. Fourthly, intra-node, the storage units are broken into many storage units to increase the bandwidth.\\

@@

\begin{figure}[h]
	\includegraphics[scale=0.3]{./images/dadiannao_node.png}
	\centering
	\caption{Block diagram of a node within Dadiannao architecture \cite{chen2014dadiannao}.}
	\label{fig:NFU_Diagram}
\end{figure}
It should be noted that although SRAMs are generally preferred over other types of memory such as DRAM and eDRAM due to the higher speed, their size is limiting factor for them. A typical eDRAM can accommodate more than $2.5\times$ data compared to a SRAM. It should be noted that more storage density is required if parameters of big models are to be stored within chip to avoid external memory access. As a result, eDRAM is selected over SRAM in Dadiannao's proposed design. Refuting the memory bandwidth limitations for for computational units allows for scaling since more inputs and outputs to each neuron can be calculated simultaneously. However, eDRAMs suffer from some drawbacks compared to SRAMs namely, slower response, periodical refreshing requirement, and destructive read that may result in inaccurate values.\\

In order to exploit high bandwidth within nodes, a tile fashion layout is implemented as shown in figure \ref{fig:tiles}. In this design output neurons are assigned to various tile units. This enables NFUs to process data from multiple neurons in parallel.
\begin{figure}[h]
	\includegraphics[scale=0.23]{./images/dadiannao_tile.png}
	\centering
	\caption{Tile implementation of a node \cite{chen2014dadiannao}.}
	\label{fig:tiles}
\end{figure}

Tiles in figure \ref{fig:tiles} are connected to eDRAM reservoirs in the middle, one for input neurons from tiles and one for output neurons. It should be noted that even though there are many tiles and NFUs in a node, the number of neurons in a typical Deep Learning architecture is much more to be a one to one relationship between NFUs and neurons. As a result, each NFU has the job of calculating many neuron outputs.

For most of the layer types, interconnect network of DaDianNao does not impose a bottleneck even despite the bulk number of neuron values, that has to be routed throughout the network. Hench, DaDianNao designer did not develop a custom-made network and adopted a 2D mesh topology with a HyperTransport 2.0 (HT2.0) IP block interconnection.

\subsection*{Experimental Results}
In this part we present the performance and energy consumption of Dadiannao. In figure \ref{fig:diannao_speedup}, performance of Dadiannao in the training mode is compared against its NVIDIA K20M GPU implementation counterpart.
\begin{figure}[h]
	\includegraphics[width=\textwidth]{./images/dadiannao_chips.png}
	\centering
	\caption{Speedup of Dadiannao/GPU for various types of layers. CONV1 and fullNN layers require at least 4 nodes. CONV3 and CONV4 need at least 36 nodes to operate\cite{chen2014dadiannao}.}
	\label{fig:dadiannao_speedup}
\end{figure}

Figure \ref{fig:dadiannao_speedup} shows that Dadiannao could reach a speedup of more than 100 over the GPU system in almost all types of layers with 16 chips or more configuration. The speedup is even higher up to more than 1000 for the CONV1 and CONV2 layers when 64 chips are adopted. It should be noted that the size of CONV3 and CONV4 layers (which have private kernels) is relatively large in the range of GBs. As a result, the performance of these layers can only be measured with more than 36-node arrangement. The FULL NN layer, which is a fully connected layer also accounts for more than 50M synaptic connections and hence cannot be implemented on a system with only 1 node. Generally, average speedup of $21.38\times,$ $79.81\times,$ $216.72\times,$and $450.65\times$ is reported for Dadiannao over GPU for 1-node, 4-node, 16-node, and 64-node configurations, respectively. However, the speedup and scaling of various types of layers are wildly different due to the inherent characteristics of each type of layer.\\

In addition, the performance of DaAianNao against GPU baseline in evaluated from energy point of view. In the training mode the energy reduction rates against the GPU baseline are recorded as $172.39\times,$ $180.42\times,$ $142.59\times,$ and $66.94\times$ for the 1-node, 4-node, 16-node, and 64-node respectively. The energy reduction number reported in the inference mode is even higher. In the inference mode, the energy reduction for the 1-node, 4-node, 16-node, and 64-node configuration is $330.56\times,$ $323.74\times,$ $270.04\times$, and $150.31\times,$ respectively.

\section*{Conclusion and Outlook}
Machine Learning and Deep Learning have already become a part of our lives that we even do not notice them anymore. From search engines, voice recognition systems in our phones, and self-driving cars, all the way to commercial and financial analytics, they all benefit from machine learning and deep learning. With the help of fast CPUs and GPUs, training of models within Machine Learning and Deep Learning frame encountered a huge boost in the recent past. However, it seems that the future attempts to guarantee the scalability of such systems should also consider designing specialized software and hardware frameworks tailored to these applications such as project Adam and DaDianNao supercomputer.\\
\subsection{High Performance Graph Processing}
Graphs are abstract mathematical concepts that have numerous applications in science and engineering. Depending on the application and context they are being adopted in, they may have different types such a directed or undirected, cyclic or acyclic, finite or infinite, etc.\\
\begin{figure}[!ht]
	\includegraphics[scale=0.25]{./images/graph_multi.png}
	\centering
	\caption{Execution mechanism of BSP programming model.}
	\label{fig:BSP}
\end{figure}
With the explosion of data over the past decade, graphs have dictated themselves as tools with many interesting properties. Nowadays, graphs are an indispensable component of large data modeling systems. Graph databases are being more and more favored over relational databases with focus on relations as connections indicate foreign-keys in the traditional relational databases. Among many interesting inherent properties of graphs, their scalability is probably the most important distinguishing factor.\\

@@

Each vertex has the option of becoming active or inactive during each iteration. If a vertex does not receive a message it can vote to halt and become inactive. Similarly, an inactive can become active upon arrival of a message. The overall execution continues until no vertex has a message to send. Figure \ref{fig:max_super} illustrates an example of finding the maximum value in a graph.

\begin{figure}[!ht]
	\includegraphics[scale=0.23]{./images/graph_superstep.png}
	\centering
	\caption{Example of computing maximum vertex value. At each superstep values for each vertex are sent to the neighbor. If the received value is higher than the current value, current value is updated. Otherwise, current value is hold \cite{sakr2013processing}.}
	\label{fig:max_super}
\end{figure}

%%%%%%% KRUNAL ADDED
\subsection{Parallel Databases and I/O}
Although the idea of parallel database system was refuted by a  \cite{dewitt1983database} DeWitt \textit{et al.} in 1983, prophesying demise of database machine, that did not stop widespread adoption of parallel database. Companies like Teradata, Tandem and many other startups pioneered a path for successful parallel database systems. The evolution to parallel databases fated due to wide adoption of relational model which favors the idea of parallel dataflow graph, leading to pipelined parallelism and partitioned parallelism. Based on the market trends, ten years later in \cite{dewitt1992parallel} DeWitt \textit{et al.} explains how parallel database system will be the future of high performance database systems. With the advances in high speed interconnect network, the client-server architecture gave much higher performance over price than its mainframe equivalent.\\

@@

Although, there is a heated argument about the performance benefits of using a Map Reduce system against the parallel DBMS system, former has been widely adopted because it offers better performance with complex queries on unstructured data. High performance computing systems have high speed network, parallel file systems and diskless nodes while Big data analytics research has evolved from an increasing demand for extracting knowledge from large data set on a big or small cluster with low speed network and each node its own disk with local file system. For example, Lustre file system is widely used in HPC environment whereas Hadoop's HDFS is one of the popular choice for big data analytic applications. The main focus of HPC systems is to optimize and exploit the resources, with a parallel file system. During the past decade, several attempts are made to bridge the gap between HPC and Big data analytics research, giving rise to new terms like High Performance Data Analysis. In the next section, we discuss about some of these research that has been propelled towards amalgamation of these two giants.\\

\subsection*{SCOPE}

Structure Computation Optimization for Parallel Execution (Scope) is a distributed computing system, that incorporates best characteristics of traditional parallel database system and map reduce execution engine, like scalability and robust implementation. In \cite{zhou2012scope}, the author describe about SCOPE's architecture and optimization techniques to achieve better performance. 
It uses SQL like query language to embrace relational query optimization techniques and separate application logic from underline implementation on a distributed system. It supports both structured and unstructured stream. All the execution plan are generated by a cost-based optimization system which is scheduled by a job scheduler, similar to MapReduce system. SCOPE runs on top of COSMOS (a distributed data platform), and used in many on-line service from Microsoft like Bing. \\

\begin{figure}[!ht]
	\includegraphics[scale=0.5]{./images/krunal8.png}
	\centering
	\caption{Architecture of COSMOS with three main components\cite{zhou2012scope}.}
	\label{fig:krunal8}
\end{figure}

\subsubsection*{COSMOS: }
Figure \ref{fig:krunal8} shows the architecture for COSMOS platform which can be divided into three main components: Storage system, Computing system and Frontend services. COSMOS storage system is a distributed append-only file system which provides fault tolerance by data replication and reliability for storing petabytes of data. Writes are always append-only and serialization is used for concurrent writes. The storage system consist of a physical partitions called extends, which is the unit of space allocation and replication, typically a size of few 100 MB. Application extends are often compressed and decompressed at client side, contains application-defined records and define a block boundary. Local or remote read and write are enabled through high-bandwidth network. Directed acyclic graph (DAG) are used for modeling a query plan. Job submission/management, job queue monitoring, job status tracking, error reporting and transferring data in and out of COSMOS are done through an interface provided by front-end services. \\

\subsubsection*{Data Representation and Query Language: }
Unstructured stream are defined as a logical sequence of bytes that requires an extractor for users to interpret it. An extractor defines the schema and iterator interface for transforming the data. It also provides an outputters that writes output in an unstructured streams. Structured stream are handled by build-in function that allow an unvarying access time by storing data with into different schema which is self-contained, that is, with rich metadata information such as schemes, structural properties and access method. Usually data is horizontally partitioned with different schemes like hash and range partitioning. Each structural stream consist of several physical extents which allows effective replication and fault tolerance. Data is stored based on store affinity and assigned affinity id, each affinity group having same id. Storage priority is given first at same machine level and if the storage is full then at same rank until its saturated and then at next close rank.  Another optimization to improve performance is stream references, in this technique the output stream is referenced with another corresponding partition based on affinity id according to storage priority. It also supports techniques like Indexing for random access, column group for vertical partitioning, etc.  \\

SCOPE query language resembles SQL and has many SQL-like extensions that enable users to separate application logic from underline implementation on a distributed system. For unstructured stream, it provides personalized commands for EXTRACT and OUTPUT, for reading from source and writing data to a sink. Structure stream removes the constraints of specifying columns because it is already contained in the stream metadata during compilation. Reference clause can be used to specify affinity of output to another stream. REDUCE command can be replaced by adding an optional PRESORT and PRODUCE clause. Figure \ref{fig:compilation_of_a_scope} shows the all steps of transformation for a SCOPE script before executing in the cluster.

\begin{figure}[!htb]
	\includegraphics[scale=0.5]{./images/krunal2}
	\centering
	\caption{Compilation of a SCOPE script.}
	\label{fig:compilation_of_a_scope}
\end{figure}

\subsubsection{Code Generation and Runtime Engine: }
SCOPE runtime engine supports all the relational operators along with build in and optimized UDOs(User Defined Opcodes) like DefaultTextExtractor and DefaultTextOutputter. Simple functions like open, next and close are performed by operator which inherently supports pipelining of these operators.  The runtime engine, after optimization, generates a physical execution tree by doing these tasks:
\begin{enumerate}
	\item The whole plan is analyzed to produce corresponding codes for operator based on the internal operator template.
	\item  Output tree is partitioned at exchanges and spool operations to generate a super vertex that contains a series of physical exchanges.
	\item Break large super vertex into smaller ones to avoid skew delay due to longer execution time.
	\item An assembly is made by compiling all the generated code into it and the entry functions for each super vertex is written into super vertex definition.
	\item Each super-vertex is schedule to execute on a single machine.
	\item Like traditional db system, a queue of additional super vertices are maintained through admission control technique when all resources are allocated.
\end{enumerate}

@@

\subsubsection{Execution Model: }
Similar to MapReduce, SCOPE implements a pull execution model. Super vertex, that reads data either locally or remotely, are executed in a pipelined fashion with an end result written to a local disk and waiting for the next vertex to pull data. This approach contrast with the execution model of parallel database which transfer intermediate data directly without any disk access. However, pull execution model has an advantage that both parties involved in data exchange need not run simultaneously and the data can be recovered from disk on failure. Since shuffling of data is the most expensive part considering the network overhead, SCOPE runtime provided many partitioning and merge operators.  It supports hash, range and index-based partitioning, with number of partitions either predetermined or determined based on intermediate data size during runtime.\\

\subsubsection{Job Scheduling: }
SCOPE's job manager is an adaption from Dryad, which takes care of job graph and resource scheduling on clusters. DAG of super vertices is scheduled to execute separately on different machines. Vertices are classified into different stages by Job manager and all the vertices in a single stage perform same task.  Data storage and computation can be performed by all the machine.
\begin{figure}[!htb]
	\includegraphics[scale=0.4]{./images/krunal3_new}
	\centering
	\caption{SCOPE job scheduling }
	\label{fig:scope_job_sched}
\end{figure}
Local system resources are maintained by processing node services and it also schedules execution of local vertices along with monitoring state of a job. Based on available resources and priorities it generates process on behalf of job managers and maintains a local queues of vertex execution request from different job manager. It stores the assembly of vertex in a cache so it can be used if the vertex is executed again. An initial DAF is formed, when the job begins, to oversee relationship among different vertex. According to execution results and variation in system environment, the job graph can dynamically adapt itself. Resource availability and vertex priority defines the actual vertex scheduling order and it's the task of job manager to notify any vertex execution failure.\\

\subsection*{Adaptive Hybrid OLAP Architecture}

Since last decade, co-processor technology has evolved rapid due to huge demand of high resolution and graphic intensive games. These co-processors are very good at processing computationally intensive tasks and since data analytics usually requires a single operation on a multiple datasets (SIMD), it's turns out to be perfect candidate. In the most of the research, the CPU ( device) acts as a supervisor and GPU (host) carries out all the task which leads to inefficient use of device resources. \\

An adaptive hybrid OLAP (On-line Analytical Processing) architecture is proposed in \cite{riha2013adaptive} that uses prediction model to efficiently distribute and process data on CPU and GPU based on their memory subsystem and memory access pattern. OLAP has a fundamental structure of a data cube which consist of multi-dimensional data that enables user to explore relation amongst different aspects of data. Data is partitioned into two types of OLAP cubes for processing on CPU and GPU:\\
\begin{enumerate}
	\item Multidimensional OLAP (MOLAP): These cubes contains preprocessed data that is sparse and huge in size. It is coherent to CPU's memory system.
	\item Relational OLAP (ROLAP): It contains data that is manipulated to appear like traditional OLAP with slicing and dicing functionality, which makes it suitable to be processed on GPU architecture.
\end{enumerate}
